---
title: "honors_thesis"
author: "Georges Ip"
date: "3/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE}
#Import libraries
library(lattice)
library(MASS)
library(car)
require(stats)
require(stats4)
library(KernSmooth)
library(cluster)
library(leaps)
library(mgcv)
library(rpart)
library(mgcv)
require("datasets")
require(graphics)
library(nlstools)
library(fpp)
library(fpp2)
library(strucchange)
library(Quandl)
library(zoo)
library(PerformanceAnalytics)
library(quantmod)
library(vars)
library(lmtest)
library(tseries)
library(dplyr)
library(lubridate)
library(TTR)
library(repmis)
library(readr)
library(SentimentAnalysis)
library(readxl)
library(tidytext)
library(wordcloud)
library(doParallel)
library(caret)
library(imputeTS)
library(randomForest)
library(gbm)
library(kernlab)
library(e1071)
library(devtools)
install.packages("remotes")
library(remotes)
remotes::install_github("trinker/sentimentr")
library(sentimentr)
library(tidyverse)
library(tm)
library(rJava)
library(coreNLP)
library(syuzhet)
library(qdap)
```

Main stock dataset
```{r}
#Using quantmod
getSymbols("^DJI", src = "yahoo", from = as.Date("2009-01-01"), to = as.Date("2019-12-31"))
DJIA<-DJI
#Add columns for returns
#frist create a name vector for new returns
names <- c("DJIA.Open.returns", "DJIA.High.returns", "DJIA.Low.returns", "DJIA.Close.returns", "DJIA.Volume.returns", "DJIA.Adjusted.returns")
returns.data<-ROC(DJIA)
names(returns.data) <- names
#Drop volume returns as it is not applicable in this case
returns.data <- subset(returns.data, select = -c(5))
#now combine the two datasets
DJIA<-merge(DJIA,returns.data,join='left') 
#Take first difference
DJIA$diff <- DJIA$DJI.Close - Lag(DJIA$DJI.Close)
#Convert difference into output vector
DJIA$direction <- ifelse(DJIA$diff  > 0, 1, 0)
DJIA$direction<-as.factor(DJIA$direction)
```

Technical Indicators (1)
```{r}
#implement a simple 10 day moving average
DJIA$SMA10 <- SMA(DJIA$DJI.Close, n = 10)
#implement a weighted 10 day moving average
DJIA$WMA10 <- WMA(DJIA$DJI.Adjusted, n = 10, wts = 1:10)
#Implement momentum, the amount the price has changed over the last n periods
momentum = diff(DJIA$DJI.Adjusted, lag = 4, differences = 1)
DJIA$momentum <- momentum
#Implement stochastic K%, another indicator comparing the closing price of an asset to a range of its prices over a certain period of time n.
#We choose n = 14 for now
DJIA$stochastic.k <- stoch(DJIA[,c("DJI.High","DJI.Low", "DJI.Adjusted")], nFastK = 14, bounded = TRUE)*100
#Implement stochastic D%, which is the 3-day mooving average of K%
DJIA$stochastic.d <- stoch(DJIA[,c("DJI.High","DJI.Low", "DJI.Adjusted")],  nFastD = 3, bounded = TRUE)*100
#Implement RSI, the relative strength index. RSI is a measure of ratio of the recent upward price movements to the absolute price movement
#We will use n=14 for now
DJIA$RSI <- RSI(DJIA$DJI.Adjusted, n = 14)
#Implement MACD, moving average convergence divergence. This is another oscillator indicator that utilizes exponential moving averages to chart thet trend of price movements 
DJIA$MACD <- MACD(DJIA[,"DJI.Close"], 12, 26, 9, maType = "EMA")
#Implement Larry William’s R%. This is a momentum indicator that tracks overbought and oversold levels. It is quite similar to the stochastic K%.
DJIA$Williams.R <- WPR(DJIA[,c("DJI.High","DJI.Low", "DJI.Close")])
#Implement the A/D (Accumulation/Distribution) Oscillator is an indicator that aims to measure bullish and bearish price pressure. 
DJIA$Williams.AD <- williamsAD(DJIA[,c("DJI.High","DJI.Low", "DJI.Close")])
#Implement the CCI (Commodity Channel Index). The CCI is another momentum indicator that is used to determine when an asset is reaching overbought or oversold conditions. 
DJIA$CCI <- CCI(DJIA[,c("DJI.High","DJI.Low", "DJI.Close")], n = 3, maType = "SMA", c = 0.015)
```

Technical Indicators (2)
```{r}
#Implement OBV, On Balance Volume. OBV is a momentum indicator that tracks changes in volume traded as well as price. 
DJIA$OBV <- OBV(DJIA[,"DJI.Close"], DJIA[,"DJI.Volume"])
#Implement a 5 period simple moving average
DJIA$SMA5 <- SMA(DJIA$DJI.Adjusted, n = 5)
#Implement a 6 period bias indicator. This indicatoor derives a directional bias relative tot he 6 period moving average value
SMA6 <- SMA(DJIA$DJI.Adjusted, n = 6)
DJIA$BIAS6 <- (DJIA$DJI.Close - SMA6)/SMA6 * 100
#Implement PSY12, which measures the ratio of rising periods over the last 12 periods
DJIA$PSY12 <- rep(NA,nrow(DJIA))
for(i in 12:nrow(DJIA)){
DJIA$PSY12[i] <- sum(DJIA$direction[(i-11):i]) /12 *100
}
#Implement ASYn, which is the average return over the past n periods
DJIA$ASY1 <- lag(DJIA$DJI.Close.returns, n=1)
DJIA$ASY2 <- SMA(DJIA$DJI.Close.returns, n =2)
DJIA$ASY3 <- SMA(DJIA$DJI.Close.returns, n =3)
DJIA$ASY4 <- SMA(DJIA$DJI.Close.returns, n =4)
DJIA$ASY5 <- SMA(DJIA$DJI.Close.returns, n =5)
```

Fundamental Indicators
```{r}
#Import the China-US exchange rate
USCHEX <- source_data("https://raw.githubusercontent.com/georgesip/financial_engineering/master/DEXCHUS-edited.csv")
USCHEX <- head(USCHEX, -1)
#Combine it with the DJIA dataset
DJIA$USCHEX <- USCHEX$DEXCHUS
#Add SP500 indicator as another fundamental
#using quantmod
sp500 <- new.env()
getSymbols("^DJI", env = sp500, src = "yahoo", from = as.Date("2009-01-01"), to = as.Date("2019-12-31"))
DJItemp <- sp500$DJI
DJI<-DJItemp

getSymbols("^DJI", src = "yahoo", from = as.Date("2009-01-01"), to = as.Date("2019-12-31"))
DJIAtemp <- ROC(DJI)
DJIA<-DJIAtemp
DJI$DJIA.returns <- DJIA$DJI.Adjusted
```

Sentiment Indicators
```{r}
#Import dataset
news1 <- read_csv("2020-3-6-15-7-23-569723201808043-S&P 500 News - Investing.com-ScrapingData-ScrapeStorm.csv")
news2 <- read_excel("2020-3-7-20-35-11-59646935038951-S&P 500 News - Investing.com-ScrapingData-ScrapeStorm-2.xlsx")
news.1 <- head(news1, -1)
news.temp <- rbind(news.1,news2)
news.full<-news.temp
```
Preprocessing of sentiment
```{r}
#Clean the dataset
date.fill <- "- Mar 05, 2020"
news.full$date[1:33] <- date.fill
news.full$date <- gsub("- ", "", news.full$date)
news.full$date <- mdy(news.full$date)
drops <- c('Title_link','js-external-link', 'articleDetails')
news.full<-news.full[,!(names(news.full) %in% drops)]
#rearrange chronological order
news.full <- arrange(news.full, date)
#There are a few duplicated entries, let's remove them
news.full<-news.full %>%
    distinct(Title, date, .keep_all = TRUE)
#remove NA's 
news.full[is.na(news.full)] = "unknown"
```


Cleaning the dataset thoroughly for SentimentAnalysis package
```{r}
#clean the text itself from headlines
news.cleaned <- news.full
news.cleaned$Title <- tolower(news.cleaned$Title)
news.cleaned$Title <- gsub("global markets-", "",news.cleaned$Title)
news.cleaned$Title <- gsub("rpt-", "",news.cleaned$Title)
news.cleaned$Title <- gsub("-", " ",news.cleaned$Title)
news.cleaned$Title <- gsub("[[:punct:]]", " ", news.cleaned$Title)
#clean the text itself from description
news.cleaned$textDiv <- tolower(news.cleaned$textDiv)
news.cleaned$textDiv <- gsub("\\s*\\([^\\)]+\\)","",news.cleaned$textDiv)
news.cleaned$textDiv <- gsub("-", " ",news.cleaned$textDiv)
news.cleaned$textDiv <- gsub("[[:punct:]]", " ", news.cleaned$textDiv)
#crete alternate set wiitout stop words
news.full.alt <- news.full
stopWords <- stopwords("SMART")
news.full.alt$Title<- as.character(news.full.alt$Title) 
'%nin%' <- Negate('%in%')
news.full.alt$Title_clean<-lapply(news.full.alt$Title, function(x) {
  chk <- unlist(strsplit(x," "))
  p <- chk[chk %nin% stopWords]
  paste(p,collapse = " ")
})
news.full.alt$desc_clean<-lapply(news.full.alt$textDiv, function(x) {
  chk <- unlist(strsplit(x," "))
  p <- chk[chk %nin% stopWords]
  paste(p,collapse = " ")
})
```

Less thorough cleaning for the other packages that can handle sentence structure better
```{r}
news.unclean <- news.full #This alternative dataset is used for the other libraries as they can handle more advanced sentence structures
#clean the text itself from headlines
news.unclean$Title <- gsub("GLOBAL MARKETS-", "",news.unclean$Title)
news.unclean$Title <- gsub("RPT-", "",news.unclean$Title)
news.unclean$Title <- gsub("-", " ",news.unclean$Title)
news.unclean$Title <- gsub("[[:punct:]]", " ", news.unclean$Title)
#clean the text itself from description
news.unclean$textDiv <- gsub("\\s*\\([^\\)]+\\)","",news.unclean$textDiv)
news.unclean$textDiv <- gsub("-", " ",news.unclean$textDiv)
news.unclean$textDiv <- gsub("[[:punct:]]", " ", news.unclean$textDiv)
news.unclean$textDiv <- gsub("[*]", ".", news.unclean$textDiv)
```


Summary of data (fine-tune this)
```{r}
#Wordcloud
wordcloud(news.cleaned$Title,min.freq = 10,colors=brewer.pal(8, "Dark2"),random.color = TRUE,max.words = 500)
Title_clean.df <- as.character(news.cleaned.alt$Title_clean)
wordcloud(Title_clean.df,min.freq = 10,colors=brewer.pal(8, "Dark2"),random.color = TRUE,max.words = 500)
```

Convert to sentiment (using SentimentAnalysis package)
```{r}
data(DictionaryLM)
#For headlines
sentiment <- analyzeSentiment(news.cleaned$Title, language = "english", aggregate = news.cleaned$date)
title.sentimentLM <- sentiment$SentimentLM
title.sentimentHE <- sentiment$SentimentHE
news.cleaned <- cbind(news.cleaned, title.sentimentLM, title.sentimentHE)
#for descriptioin as a sanity check
sentiment2 <- analyzeSentiment(news.cleaned$textDiv, language = "english", aggregate = news.cleaned$date)
desc.sentimentLM <- sentiment2$SentimentLM
desc.sentimentHE <- sentiment2$SentimentHE
news.cleaned <- cbind(news.cleaned, desc.sentimentLM, desc.sentimentHE)
#Do the same for the alternate sentiment set
sentiment.alt <- analyzeSentiment(news.cleaned.alt$Title, language = "english", aggregate = news.cleaned.alt$date)
title.sentimentLM.alt <- sentiment.alt$SentimentLM
title.sentimentHE.alt <- sentiment.alt$SentimentHE
news.cleaned.alt <- cbind(news.cleaned.alt, title.sentimentLM.alt, title.sentimentHE.alt)
#for description as a sanity check
sentiment2.alt <- analyzeSentiment(news.cleaned.alt$textDiv, language = "english", aggregate = news.cleaned.alt$date)
desc.sentimentLM.alt <- sentiment2.alt$SentimentLM
desc.sentimentHE.alt <- sentiment2.alt$SentimentHE
news.cleaned.alt <- cbind(news.cleaned.alt, desc.sentimentLM.alt, desc.sentimentHE.alt)
```

It seems that when we run the sentiment analysis on the description, we obtain greater granular detail. Whereas the short length of the headlines results in multiple zero entries. (For the sentimentanalysis package).


```{r}
#We can also use the other packages to compare performance
#Use sentimentr package, which is an augmented dictionary lookup that takes valence shifters into account.
#work on the titles first
out.title <- with(
    news.unclean, 
    sentiment_by(
        get_sentences(Title), 
        date
    )
)
sentiment.ts <- xts(out$ave_sentiment, order.by=out.title$date)
plot(sentiment.ts)
#Now work on descriptions
(out.desc <- with(
    news.unclean, 
    sentiment_by(
        get_sentences(textDiv), 
        date
    )
))
sentiment.desc.ts <- xts(out.desc$ave_sentiment, order.by=out.desc$date)
plot(sentiment.desc.ts)
#Another possible plot
highlight(out.title)
highlight(out.desc)
```


Syuzhet package
```{r}
#Another option is the syuzhet package

#Using Stanford coreNLP in syuzhet

```


Using stanford coreNLP (doesn't work, ignore for now)
```{r}
volatiles = new.env(parent=emptyenv())
tagger_path <- "C:\\Users\\Georges\\Desktop\\stanfordNLP"
sentences.desc <- get_sentences(news.unclean$textDiv)
stanford.annotate <- annotateString(sentences.desc)
sentiment.stanford <- get_sentiment(sentences.desc, method="stanford", tagger_path)
```


Integrate sentiment data with DJI (SentimentAnalysis)
```{r}
#First, group by date
news.sentiment.title.LM <- aggregate(news.cleaned["title.sentimentLM"], by=news.full["date"], mean)
news.sentiment.desc.LM <- aggregate(news.cleaned["desc.sentimentLM"], by=news.full["date"], mean)
news.sentiment.title.HE <- aggregate(news.cleaned["title.sentimentHE"], by=news.full["date"], mean)
news.sentiment.desc.HE <- aggregate(news.cleaned["desc.sentimentHE"], by=news.full["date"], mean)
```


Match SentimentAnalysis to DJI dataset based on date
```{r}
title.sa.LM <- rep(NA,nrow(DJI))
title.sa.HE <- rep(NA,nrow(DJI))
desc.sa.LM <- rep(NA,nrow(DJI))
desc.sa.HE <- rep(NA,nrow(DJI))

#Start with LM
for(i in 1:nrow(DJI)){
  if(any(as.Date.character(index(DJI)[i])==news.sentiment.title.LM$date)){
         title.sa.LM[i] <- news.sentiment.title.LM$title.sentimentLM[which(as.Date.character(index(DJI)[i])==news.sentiment.title.LM$date)]
  }
  if(any(as.Date.character(index(DJI)[i])==news.sentiment.desc.LM$date)){
         desc.sa.LM[i] <- news.sentiment.desc.LM$desc.sentimentLM[which(as.Date.character(index(DJI)[i])==news.sentiment.desc.LM$date)]
  }
}
DJI$title.sa <- title.sa.LM
DJI$desc.sa <- desc.sa.LM
#now with HE
for(i in 1:nrow(DJI)){
  if(any(as.Date.character(index(DJI)[i])==news.sentiment.title.HE$date)){
         title.sa.HE[i] <- news.sentiment.title.HE$title.sentimentHE[which(as.Date.character(index(DJI)[i])==news.sentiment.title.HE$date)]
  }
  if(any(as.Date.character(index(DJI)[i])==news.sentiment.desc.HE$date)){
         desc.sa.HE[i] <- news.sentiment.desc.HE$desc.sentimentHE[which(as.Date.character(index(DJI)[i])==news.sentiment.desc.HE$date)]
  }
}
DJI$title.sa <- title.sa.HE
DJI$desc.sa <- desc.sa.HE
```


Implement sentiment from sentimentr into the dataset
```{r}
title.sr <- rep(NA,nrow(DJI))
desc.sr <- rep(NA,nrow(DJI))

for(i in 1:nrow(DJI)){
  if(any(as.Date.character(index(DJI)[i])==out.title$date)){
         title.sr[i] <- out.title$ave_sentiment[which(as.Date.character(index(DJI)[i])==out.title$date)]
  }
  if(any(as.Date.character(index(DJI)[i])==out.desc$date)){
         desc.sr[i] <- out.desc$ave_sentiment[which(as.Date.character(index(DJI)[i])==out.desc$date)]
  }
  
}
DJI$title.sr <- title.sr
DJI$desc.sr <- desc.sr
```


We have to offset our predictors as they are ordered time series. Hence, we aim to use predictors calculated at time t,to predict the direction at tme t+1.
```{r}
full.DJI <- DJI
#lag predictors by 1 day
for (i in 14:43){
  DJI[,i] <- lag(DJI[,i],1)
}
```


We now look at plots of the data
```{r}
plot.zoo(DJI$diff, xlab = "Month", ylab = "First difference of adjusted returns", col = 'blue', lwd = 0.5)
```
We can see that taking the first difference detrended the time series, and cetered it around zero. However, variance is still not constant, there is also clear signs of volatility clustering. However, we are focused on the direction of the time series, thus, the stationarity requirement is not as stringent. 



We then perform a test for granger causality between the difference and our different fundamental and sentiment variables
```{r}
#For DJIA with DJI
grangertest(diff~DJIA.returns, order = 1, data=DJI)
grangertest(DJIA.returns~diff, order = 1, data=DJI)
```
It is evident that for sentiment scores constructed from headlines, the difference in prices granger causes sentiment.


```{r}
grangertest(diff~DJI_sentiment_desc, order = 1, data=DJI)
grangertest(DJI_sentiment_desc~diff, order = 1, data=DJI)
```



We cannot perform the usual resampling methods such as LOOCV or k-fold CV as our data is ordered by time. However, the caret library allows for this by using a rolling window estimation method. 
```{r}
#We have to first preprocess the data, as there are huge gaps in the news dataset.
#subset DJI (missing news before 2014-03-25)
DJI <- DJI[1315:2767,]
#convert to dataframe for methods that require it
set.seed(123)
DJI.clean.df <- fortify.zoo(DJI)
DJI.clean.df$direction <- factor(as.numeric(DJI.clean.df$direction == 2))
rownames(DJI.clean.df) <- DJI.clean.df[,1]
DJI.clean.df <- DJI.clean.df[,-1]
levels(DJI.clean.df$direction) <- c("down", "up")
DJI.clean.df$direction <- as.factor(DJI.clean.df$direction)
#Remove all non-relevant variables from DJI
drops <- c("DJI.Open","DJI.High","DJI.Low","DJI.Close","DJI.Volume","DJI.Adjusted","DJI.Open.returns","DJI.High.returns", "DJI.Low.returns","DJI.Close.returns", "DJI.Adjusted.returns", "diff")
DJI.clean<-DJI.clean.df[ , !(names(DJI.clean.df) %in% drops)]
```


The first dataset will be the full one with all variables included
```{r}
#Assess the number of NA's
colSums(is.na(DJI.clean))
#There are still some NA's in USCHEX, we can do some imputation
DJI.clean <- na.locf(DJI.clean)
#Split into training and test set
DJI.end <- floor(0.8*nrow(DJI.clean)) #select the first 80% of the data
DJI.train <- DJI.clean[1:DJI.end,] #assign the first 80% of the data to the train set
DJI.test <- DJI.clean[(DJI.end+1):nrow(DJI.clean),]
```


Create dataset for technical indicators (1) only, without sentiment and fundamentals
```{r}
DJI.1 <- DJI.clean
drops1 <- c("OBV", "SMA5","BIAS6","PSY12","ASY1", "ASY2","ASY3","ASY4","ASY5","USCHEX","DJIA.returns", "DJI_sentiment_desc")
DJI.1<-DJI.1[ , !(names(DJI.1) %in% drops3)]
#Split into training and test set
DJI.train.1 <- DJI.1[1:DJI.end,] #assign the first 80% of the data to the train set
DJI.test.1 <- DJI.1[(DJI.end+1):nrow(DJI.1),]
```



Create dataset for technical indicators (2) only, without sentiment and fundamentals
```{r}
DJI.2 <- DJI.clean
drops2 <- c("SMA10", "WMA10","momentum","fastK","fastD","stochastic.k","fastK.1","fastD.1","stochastic.d","RSI"      ,"macd","MACD","Williams.R","Williams.AD","CCI","USCHEX","DJIA.returns","DJI_sentiment_title", "DJI_sentiment_desc")
DJI.2<-DJI.2[ , !(names(DJI.2) %in% drops2)]
#Split into training and test set
DJI.train.2 <- DJI.2[1:DJI.end,] #assign the first 80% of the data to the train set
DJI.test.2 <- DJI.2[(DJI.end+1):nrow(DJI.2),]
```


Create dataset with fundamentals only
```{r}
DJI.f <- DJI.clean
dropsf <- c("SMA10", "WMA10","momentum","fastK","fastD","stochastic.k","fastK.1","fastD.1","stochastic.d","RSI"      ,"macd","MACD","Williams.R","Williams.AD","CCI","OBV", "SMA5","BIAS6","PSY12","ASY1", "ASY2","ASY3","ASY4","ASY5","DJI_sentiment_title", "DJI_sentiment_desc")
DJI.f<-DJI.f[ , !(names(DJI.f) %in% dropsf)]
#Split into training and test set
DJI.train.f <- DJI.f[1:DJI.end,] #assign the first 80% of the data to the train set
DJI.test.f <- DJI.2[(DJI.end+1):nrow(DJI.f),]
```

Create dataset with only sentiment
```{r}
DJI.s <- DJI.clean
dropss <- c("DJI_sentiment_title", "DJI_sentiment_desc")
DJI.s<-DJI.s[ , !(names(DJI.s) %in% dropss)]
#Split into training and test set
DJI.train.s <- DJI.s[1:DJI.end,] #assign the first 80% of the data to the train set
DJI.test.s <- DJI.s[(DJI.end+1):nrow(DJI.s),]
```


Prepare for caret library
```{r}
set.seed(1)
#prepare parallel computing method
registerDoParallel(cores=6)
myTimeControl <- trainControl(method = "timeslice",
                              initialWindow = 32,
                              horizon = 2,
                              fixedWindow = FALSE,
                              classProbs = TRUE, 
                              summaryFunction = twoClassSummary,
                              allowParallel = TRUE)
```


LASSO/RIDGE with technicals (1), fundamentals, and sentiment desc
```{r}
#optimise using AIC and BIC

lasso.grid <- expand.grid(lambda=seq(0,1,by=0.05),alpha=1)
lasso.fit <- train(direction ~ .,
                    data = DJI.train,
                    method = "glmnet",
                    metric = "ROC",
                    family = "binomial",
                    preProcess = c("range"),
                    trControl = myTimeControl,
                    tuneGrid = lasso.grid)
lasso.pred <- predict(lasso.fit, DJI.test)
confusionMatrix(lasso.pred, DJI.test$direction)
mean(lasso.pred == DJI.test$direction)                    
```
We obtain an accuracy of only 46.7%


Lasso with technical (1) only

```{r}
lasso.fit1 <- train(direction ~ .,
                    data = DJI.train1,
                    method = "glmnet",
                    metric = "ROC",
                    family = "binomial",
                    preProcess = c("range"),
                    trControl = myTimeControl,
                    tuneGrid = lasso.grid)
lasso.pred2 <- predict(lasso.fit1, DJI.test1)
confusionMatrix(lasso.pred2, DJI.test2$direction)
```
We obtain an accuracy of 49.48%


Lasso with technical (2) only
```{r}
lasso.fit2 <- train(direction ~ .,
                    data = DJI.train2,
                    method = "glmnet",
                    metric = "ROC",
                    family = "binomial",
                    preProcess = c("range"),
                    trControl = myTimeControl,
                    tuneGrid = lasso.grid)
lasso.pred2 <- predict(lasso.fit2, DJI.test2)
confusionMatrix(lasso.pred2, DJI.test2$direction)
mean(lasso.pred2 == DJI.test2$direction)    
```
We obtain an accuracy of 57.39%


Lasso with fundamentals only
```{r}
lasso.fit.f <- train(direction ~ .,
                    data = DJI.train.f,
                    method = "glmnet",
                    metric = "ROC",
                    family = "binomial",
                    preProcess = c("range"),
                    trControl = myTimeControl,
                    tuneGrid = lasso.grid)
lasso.pred.f <- predict(lasso.fit.f, DJI.test.f)
confusionMatrix(lasso.pred.f, DJI.test.f$direction)
```

LASSO wiith sentiment only from SentimentAnalysis method
```{r}

```

LASSO with sentiment only from sentimentr method
```{r}

```


Random Forest
```{r}
#a quick way to find a good value for mtry
#bestMtry <- tuneRF(DJI.train[,2:ncol(DJI.train)], DJI.train$direction, stepFactor = 1.5, improve = 1e-5, ntree = 1000)
#Using caret to grid search
rf.control <-trainControl(method = "timeslice",
                              initialWindow = 32,
                              horizon = 12,
                              fixedWindow = FALSE,
                              number = 10,
                              classProbs = TRUE, 
                              allowParallel = TRUE)
mtry <- sqrt(ncol(DJI.train))
rfgrid <- expand.grid(.mtry=mtry)
rf.fit <- train(direction ~ .,
                    data = DJI.train,
                    method = "rf",
                    metric = 'Accuracy',
                    preProcess = c("range"),
                    trControl = rf.control,
                    tuneGrid = rfgrid)
print(rf.fit)
rf.test <- predict(rf.fit, newdata = DJI.test)
confusionMatrix(rf.test, DJI.test$direction)
#Alternatively
tune_grid <- expand.grid(
  mtry = c(
    ncol(DJI.train)-1, # p
    ncol((DJI.train)-1)/ 3, # p / 3
    ceiling(sqrt(ncol(DJI.train))) # square root of p
  )
)
rf.control2 <-trainControl(method = "timeslice",
                              initialWindow = 32,
                              horizon = 12,
                              fixedWindow = FALSE,
                              number = 10,
                              search = "random",
                              classProbs = TRUE, 
                              allowParallel = TRUE,
                              repeats = 3)
rf.fit2 <- train(direction ~ .,
                    data = DJI.train,
                    method = "rf",
                    metric = 'Accuracy',
                    preProcess = c("range"),
                    trControl = rf.control2,
                    tuneGrid = tune_grid)
print(rf.fit2)
rf.test2 <- predict(rf.fit2, newdata = DJI.test)
confusionMatrix(rf.test2, DJI.test$direction)
```


SVM
```{r}
#First we find the optimal tuning parameter
svm.tune <- train(direction~.,
                  data = DJI.train,
                  method = "svmRadial",   # Radial kernel
                  tuneLength = 9,					# 9 values of the cost function
                  preProc = c("range"),  # Center and scale data
                  metric="ROC",
                  trControl=myTimeControl)
svm.tune
#we narrow the grid search
svm.grid <- expand.grid(sigma = c(.040, .036, 0.030), C = c(0.25, 0.2, 0.35))
svm.tune2 <- train(direction~.,
                  data = DJI.train,
                  method = "svmRadial",   # Radial kernel
                  tuneGrid = svm.grid,					# 9 values of the cost function
                  preProc = c("range"),  # Center and scale data
                  metric="ROC",
                  trControl=myTimeControl)
svm.test <- predict(svm.tune, newdata = DJI.test)
confusionMatrix(svm.test, DJI.test$direction)
```
Using technical indicators 2
```{r}
svm.tune3 <- train(direction~.,
                  data = DJI.train2,
                  method = "svmRadial",   # Radial kernel
                  tuneLength = 9,					# 9 values of the cost function
                  preProc = c("range"),  # Center and scale data
                  metric="ROC",
                  trControl=myTimeControl)
svm.tune2
#we narrow the grid search
svm.grid <- expand.grid(sigma = c(.040, .036, 0.030), C = c(0.25, 0.2, 0.35))
svm.tune4 <- train(direction~.,
                  data = DJI.train2,
                  method = "svmRadial",   # Radial kernel
                  tuneGrid = svm.grid,					# 9 values of the cost function
                  preProc = c("range"),  # Center and scale data
                  metric="ROC",
                  trControl=myTimeControl)
svm.test3 <- predict(svm.tune2, newdata = DJI.test2)
confusionMatrix(svm.test3, DJI.test2$direction)
```



