---
title: "honors_thesis"
author: "Georges Ip"
date: "3/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE}
#Import libraries
library(lattice)
library(MASS)
library(car)
require(stats)
require(stats4)
library(KernSmooth)
library(cluster)
library(leaps)
library(mgcv)
library(rpart)
library(mgcv)
require("datasets")
require(graphics)
library(nlstools)
library(fpp)
library(fpp2)
library(strucchange)
library(Quandl)
library(zoo)
library(PerformanceAnalytics)
library(quantmod)
library(vars)
library(lmtest)
library(tseries)
library(dplyr)
library(lubridate)
library(TTR)
library(repmis)
library(readr)
library(SentimentAnalysis)
library(readxl)
library(tidytext)
library(wordcloud)
library(doParallel)
library(caret)
library(imputeTS)
library(randomForest)
library(gbm)
library(kernlab)
library(e1071)
library(devtools)
install.packages("remotes")
library(remotes)
remotes::install_github("trinker/sentimentr")
library(sentimentr)
library(tidyverse)
library(tm)
library(rJava)
library(coreNLP)
library(syuzhet)
library(qdap)
library(caretEnsemble)
```

Main stock dataset
```{r message=FALSE}
#using quantmod
getSymbols("^GSPC", src = "yahoo", from = as.Date("2009-01-01"), to = as.Date("2019-12-31"))
```

```{r}
#Add columns for returns
#frist create a name vector for new returns
names <- c("GSPC.Open.returns", "GSPC.High.returns", "GSPC.Low.returns", "GSPC.Close.returns", "GSPC.Volume.returns", "GSPC.Adjusted.returns")
returns.data<-ROC(GSPC)
names(returns.data) <- names
#Drop volume returns as it is not applicable in this case
returns.data <- subset(returns.data, select = -c(5))
#now combine the two datasets
GSPC<-merge(GSPC,returns.data,join='left') 
#Take first difference
GSPC$diff <- GSPC$GSPC.Adjusted - Lag(GSPC$GSPC.Adjusted)
#Convert difference into output vector
GSPC$direction <- ifelse(GSPC$diff  > 0, 1, 0)
GSPC$direction<-as.factor(GSPC$direction)
```

Technical Indicators (1)
```{r}
#implement a simple 10 day moving average
GSPC$SMA10 <- SMA(GSPC$GSPC.Close, n = 10)
#implement a weighted 10 day moving average
GSPC$WMA10 <- WMA(GSPC$GSPC.Adjusted, n = 10, wts = 1:10)
#Implement momentum, the amount the price has changed over the last n periods
momentum = diff(GSPC$GSPC.Adjusted, lag = 4, differences = 1)
GSPC$momentum <- momentum
#Implement stochastic K%, another indicator comparing the closing price of an asset to a range of its prices over a certain period of time n.
#We choose n = 14 for now
GSPC$stochastic.k <- stoch(GSPC[,c("GSPC.High","GSPC.Low", "GSPC.Adjusted")], nFastK = 14, bounded = TRUE)*100
#Implement stochastic D%, which is the 3-day mooving average of K%
GSPC$stochastic.d <- stoch(GSPC[,c("GSPC.High","GSPC.Low", "GSPC.Adjusted")],  nFastD = 3, bounded = TRUE)*100
#Implement RSI, the relative strength index. RSI is a measure of ratio of the recent upward price movements to the absolute price movement
#We will use n=14 for now
GSPC$RSI <- RSI(GSPC$GSPC.Adjusted, n = 14)
#Implement MACD, moving average convergence divergence. This is another oscillator indicator that utilizes exponential moving averages to chart thet trend of price movements 
GSPC$MACD <- MACD(GSPC[,"GSPC.Close"], 12, 26, 9, maType = "EMA")
#Implement Larry William’s R%. This is a momentum indicator that tracks overbought and oversold levels. It is quite similar to the stochastic K%.
GSPC$Williams.R <- WPR(GSPC[,c("GSPC.High","GSPC.Low", "GSPC.Close")])
#Implement the A/D (Accumulation/Distribution) Oscillator is an indicator that aims to measure bullish and bearish price pressure. 
GSPC$Williams.AD <- williamsAD(GSPC[,c("GSPC.High","GSPC.Low", "GSPC.Close")])
#Implement the CCI (Commodity Channel Index). The CCI is another momentum indicator that is used to determine when an asset is reaching overbought or oversold conditions. 
GSPC$CCI <- CCI(GSPC[,c("GSPC.High","GSPC.Low", "GSPC.Close")], n = 3, maType = "SMA", c = 0.015)
```

Technical Indicators (2)
```{r}

#Implement OBV, On Balance Volume. OBV is a momentum indicator that tracks changes in volume traded as well as price. 
GSPC$OBV <- OBV(GSPC[,"GSPC.Close"], GSPC[,"GSPC.Volume"])
#Implement a 5 period simple moving average
GSPC$SMA5 <- SMA(GSPC$GSPC.Adjusted, n = 5)
#Implement a 6 period bias indicator. This indicatoor derives a directional bias relative tot he 6 period moving average value
SMA6 <- SMA(GSPC$GSPC.Adjusted, n = 6)
GSPC$BIAS6 <- (GSPC$GSPC.Close - SMA6)/SMA6 * 100
#Implement PSY12, which measures the ratio of rising periods over the last 12 periods

GSPC$PSY12 <- rep(NA,nrow(GSPC))
for(i in 12:nrow(GSPC)){
GSPC$PSY12[i] <- sum(GSPC$direction[(i-11):i]) /12 *100
}
#Create a dummy variable that counts the number of 1's in the last 12 periods
psy.count <- GSPC$direction
#GSPC$PSY12 <- 
#Implement ASYn, which is the average return over the past n periods
GSPC$ASY1 <- lag(GSPC$GSPC.Close.returns, n=1)
GSPC$ASY2 <- SMA(GSPC$GSPC.Close.returns, n =2)
GSPC$ASY3 <- SMA(GSPC$GSPC.Close.returns, n =3)
GSPC$ASY4 <- SMA(GSPC$GSPC.Close.returns, n =4)
GSPC$ASY5 <- SMA(GSPC$GSPC.Close.returns, n =5)
```

Fundamental Indicators
```{r}
#Import the China-US exchange rate
USCHEX <- source_data("https://raw.githubusercontent.com/georgesip/financial_engineering/master/DEXCHUS-edited.csv")
USCHEX <- head(USCHEX, -1)
#Combine it with the GSPC dataset
GSPC$USCHEX <- USCHEX$DEXCHUS
#Add DJIA indicator as another fundamental
getSymbols("^DJI", src = "yahoo", from = as.Date("2009-01-01"), to = as.Date("2019-12-31"))
DJIA_returns <- ROC(DJI)
GSPC$DJIA.returns <- DJIA_returns$DJI.Adjusted
```

Sentiment Indicators
```{r}
#Import dataset
news1 <- read_csv("2020-3-6-15-7-23-569723201808043-S&P 500 News - Investing.com-ScrapingData-ScrapeStorm.csv")
news2 <- read_excel("2020-3-7-20-35-11-59646935038951-S&P 500 News - Investing.com-ScrapingData-ScrapeStorm-2.xlsx")
news.1 <- head(news1, -1)
news.temp <- rbind(news.1,news2)
news.full<-news.temp
```
Preprocessing of sentiment
```{r}
#Clean the dataset
date.fill <- "- Mar 05, 2020"
news.full$date[1:33] <- date.fill
news.full$date <- gsub("- ", "", news.full$date)
news.full$date <- mdy(news.full$date)
drops <- c('Title_link','js-external-link', 'articleDetails')
news.full<-news.full[,!(names(news.full) %in% drops)]
#rearrange chronological order
news.full <- arrange(news.full, date)
#There are a few duplicated entries, let's remove them
news.full<-news.full %>%
    distinct(Title, date, .keep_all = TRUE)
#remove NA's 
news.full[is.na(news.full)] = "unknown"
```


Cleaning the dataset thoroughly for SentimentAnalysis package
```{r}
#clean the text itself from headlines
news.cleaned <- news.full
news.cleaned$Title <- tolower(news.cleaned$Title)
news.cleaned$Title <- gsub("global markets-", "",news.cleaned$Title)
news.cleaned$Title <- gsub("rpt-", "",news.cleaned$Title)
news.cleaned$Title <- gsub("-", " ",news.cleaned$Title)
news.cleaned$Title <- gsub("[[:punct:]]", " ", news.cleaned$Title)
#clean the text itself from description
news.cleaned$textDiv <- tolower(news.cleaned$textDiv)
news.cleaned$textDiv <- gsub("\\s*\\([^\\)]+\\)","",news.cleaned$textDiv)
news.cleaned$textDiv <- gsub("-", " ",news.cleaned$textDiv)
news.cleaned$textDiv <- gsub("[[:punct:]]", " ", news.cleaned$textDiv)
#crete alternate set wiitout stop words
news.full.alt <- news.full
stopWords <- stopwords("SMART")
news.full.alt$Title<- as.character(news.full.alt$Title) 
'%nin%' <- Negate('%in%')
news.full.alt$Title_clean<-lapply(news.full.alt$Title, function(x) {
  chk <- unlist(strsplit(x," "))
  p <- chk[chk %nin% stopWords]
  paste(p,collapse = " ")
})
news.full.alt$desc_clean<-lapply(news.full.alt$textDiv, function(x) {
  chk <- unlist(strsplit(x," "))
  p <- chk[chk %nin% stopWords]
  paste(p,collapse = " ")
})
```

Less thorough cleaning for the other packages that can handle sentence structure better
```{r}
news.unclean <- news.full #This alternative dataset is used for the other libraries as they can handle more advanced sentence structures
#clean the text itself from headlines
news.unclean$Title <- gsub("GLOBAL MARKETS-", "",news.unclean$Title)
news.unclean$Title <- gsub("RPT-", "",news.unclean$Title)
news.unclean$Title <- gsub("-", " ",news.unclean$Title)
news.unclean$Title <- gsub("[[:punct:]]", " ", news.unclean$Title)
#clean the text itself from description
news.unclean$textDiv <- gsub("\\s*\\([^\\)]+\\)","",news.unclean$textDiv)
news.unclean$textDiv <- gsub("-", " ",news.unclean$textDiv)
news.unclean$textDiv <- gsub("[[:punct:]]", " ", news.unclean$textDiv)
news.unclean$textDiv <- gsub("[*]", ".", news.unclean$textDiv)
```


Summary of data (fine-tune this)
```{r}
#Wordcloud
wordcloud(news.cleaned$Title,min.freq = 10,colors=brewer.pal(8, "Dark2"),random.color = TRUE,max.words = 500)
Title_clean.df <- as.character(news.cleaned.alt$Title_clean)
wordcloud(Title_clean.df,min.freq = 10,colors=brewer.pal(8, "Dark2"),random.color = TRUE,max.words = 500)
```

Convert to sentiment (using SentimentAnalysis package)
```{r}
data(DictionaryLM)
#For headlines
sentiment <- analyzeSentiment(news.cleaned$Title, language = "english", aggregate = news.cleaned$date)
title.sentimentLM <- sentiment$SentimentLM
title.sentimentHE <- sentiment$SentimentHE
news.cleaned <- cbind(news.cleaned, title.sentimentLM, title.sentimentHE)
#for descriptioin as a sanity check
sentiment2 <- analyzeSentiment(news.cleaned$textDiv, language = "english", aggregate = news.cleaned$date)
desc.sentimentLM <- sentiment2$SentimentLM
desc.sentimentHE <- sentiment2$SentimentHE
news.cleaned <- cbind(news.cleaned, desc.sentimentLM, desc.sentimentHE)
#Do the same for the alternate sentiment set
sentiment.alt <- analyzeSentiment(news.cleaned.alt$Title, language = "english", aggregate = news.cleaned.alt$date)
title.sentimentLM.alt <- sentiment.alt$SentimentLM
title.sentimentHE.alt <- sentiment.alt$SentimentHE
news.cleaned.alt <- cbind(news.cleaned.alt, title.sentimentLM.alt, title.sentimentHE.alt)
#for description as a sanity check
sentiment2.alt <- analyzeSentiment(news.cleaned.alt$textDiv, language = "english", aggregate = news.cleaned.alt$date)
desc.sentimentLM.alt <- sentiment2.alt$SentimentLM
desc.sentimentHE.alt <- sentiment2.alt$SentimentHE
news.cleaned.alt <- cbind(news.cleaned.alt, desc.sentimentLM.alt, desc.sentimentHE.alt)
```

It seems that when we run the sentiment analysis on the description, we obtain greater granular detail. Whereas the short length of the headlines results in multiple zero entries. (For the sentimentanalysis package).


```{r}
#We can also use the other packages to compare performance
#Use sentimentr package, which is an augmented dictionary lookup that takes valence shifters into account.
#work on the titles first
out.title <- with(
    news.unclean, 
    sentiment_by(
        get_sentences(Title), 
        date
    )
)
sentiment.ts <- xts(out$ave_sentiment, order.by=out.title$date)
plot(sentiment.ts)
#Now work on descriptions
(out.desc <- with(
    news.unclean, 
    sentiment_by(
        get_sentences(textDiv), 
        date
    )
))
sentiment.desc.ts <- xts(out.desc$ave_sentiment, order.by=out.desc$date)
plot(sentiment.desc.ts)
#Another possible plot
highlight(out.title)
highlight(out.desc)
```


Syuzhet package
```{r}
#Another option is the syuzhet package

#Using Stanford coreNLP in syuzhet

```


Using stanford coreNLP (doesn't work, ignore for now)
```{r}
#volatiles = new.env(parent=emptyenv())
#tagger_path <- "C:\\Users\\Georges\\Desktop\\stanfordNLP"
#sentences.desc <- get_sentences(news.unclean$textDiv)
#stanford.annotate <- annotateString(sentences.desc)
#sentiment.stanford <- get_sentiment(sentences.desc, method="stanford", tagger_path)
```


Integrate sentiment data with GSPC (SentimentAnalysis)
```{r}
#First, group by date
news.sentiment.title.LM <- aggregate(news.cleaned["title.sentimentLM"], by=news.cleaned["date"], mean)
news.sentiment.desc.LM <- aggregate(news.cleaned["desc.sentimentLM"], by=news.cleaned["date"], mean)
news.sentiment.title.HE <- aggregate(news.cleaned["title.sentimentHE"], by=news.cleaned["date"], mean)
news.sentiment.desc.HE <- aggregate(news.cleaned["desc.sentimentHE"], by=news.cleaned["date"], mean)
```


Match SentimentAnalysis to GSPC dataset based on date
```{r}
title.sa.LM <- rep(NA,nrow(GSPC))
title.sa.HE <- rep(NA,nrow(GSPC))
desc.sa.LM <- rep(NA,nrow(GSPC))
desc.sa.HE <- rep(NA,nrow(GSPC))

#Start with LM
for(i in 1:nrow(GSPC)){
  if(any(as.Date.character(index(GSPC)[i])==news.sentiment.title.LM$date)){
         title.sa.LM[i] <- news.sentiment.title.LM$title.sentimentLM[which(as.Date.character(index(GSPC)[i])==news.sentiment.title.LM$date)]
  }
  if(any(as.Date.character(index(GSPC)[i])==news.sentiment.desc.LM$date)){
         desc.sa.LM[i] <- news.sentiment.desc.LM$desc.sentimentLM[which(as.Date.character(index(GSPC)[i])==news.sentiment.desc.LM$date)]
  }
}
GSPC$title.sa.LM <- title.sa.LM
GSPC$desc.sa.LM <- desc.sa.LM
#now with HE
for(i in 1:nrow(GSPC)){
  if(any(as.Date.character(index(GSPC)[i])==news.sentiment.title.HE$date)){
         title.sa.HE[i] <- news.sentiment.title.HE$title.sentimentHE[which(as.Date.character(index(GSPC)[i])==news.sentiment.title.HE$date)]
  }
  if(any(as.Date.character(index(GSPC)[i])==news.sentiment.desc.HE$date)){
         desc.sa.HE[i] <- news.sentiment.desc.HE$desc.sentimentHE[which(as.Date.character(index(GSPC)[i])==news.sentiment.desc.HE$date)]
  }
}
GSPC$title.sa.HE <- title.sa.HE
GSPC$desc.sa.HE <- desc.sa.HE
```


Implement sentiment from sentimentr into the dataset
```{r}
title.sr <- rep(NA,nrow(GSPC))
desc.sr <- rep(NA,nrow(GSPC))

for(i in 1:nrow(GSPC)){
  if(any(as.Date.character(index(GSPC)[i])==out.title$date)){
         title.sr[i] <- out.title$ave_sentiment[which(as.Date.character(index(GSPC)[i])==out.title$date)]
  }
  if(any(as.Date.character(index(GSPC)[i])==out.desc$date)){
         desc.sr[i] <- out.desc$ave_sentiment[which(as.Date.character(index(GSPC)[i])==out.desc$date)]
  }
  
}
GSPC$title.sr <- title.sr
GSPC$desc.sr <- desc.sr
```

There are huge gaps in the news dataset, identify the longest contiguous stretch without NAs and subset the data
```{r}
#We first create another dataset with all the sentiment scores from all methods
keeps <- c("title.sa.LM","desc.sa.LM", "title.sa.HE", "desc.sa.HE", "title.sr","desc.sr")
sentiment_scores<-GSPC[ , (names(GSPC) %in% keeps)]
#Find the longest stretch
stretch <- na.contiguous(sentiment_scores)
#subset GSPC (missing news before 2014-03-25)
GSPC_full <- GSPC
GSPC <- GSPC[1315:2767,]

#There are still some NA's in USCHEX, we can do some imputation
colSums(is.na(GSPC))
GSPC <- na_locf(GSPC)
```


We now look at plots of the data
```{r}
#Time series
plot.zoo(GSPC$diff, xlab = "Month", ylab = "First difference of adjusted returns", col = 'blue', lwd = 0.5)
#Correlation heatmap
corr.df <- subset(GSPC, select = -c(GSPC.Open,GSPC.High,GSPC.Low,GSPC.Close,GSPC.Volume,GSPC.Adjusted,GSPC.Open.returns,GSPC.High.returns, GSPC.Low.returns,GSPC.Adjusted.returns))
corr <- round(cor(corr.df),1)
library(ggcorrplot)
ggcorrplot(corr, tl.cex = 3)
```
We can see that taking the first difference detrended the time series, and cetered it around zero. However, variance is still not constant, there is also clear signs of volatility clustering. However, we are focused on the direction of the time series, thus, the stationarity requirement is not as stringent. 

The corrplot gives us a rough idea of which of our variables might end up being useful.


We then perform a test for granger causality between the difference and our different fundamental and sentiment variables
```{r}
#For DJIA with GSPC
keeps <- c("GSPC.Close.returns","DJIA.returns")
GSPC.granger.1 <-GSPC[ ,(names(GSPC) %in% keeps)]
VARselect(GSPC.granger.1, lag.max = 7)
grangertest(GSPC.Close.returns~DJIA.returns, order = 1, data=GSPC.granger.1)
```

```{r}
#For GSPC with USCHEX
keeps2 <- c("GSPC.Close.returns","USCHEX")
GSPC.granger.2 <-GSPC[ ,(names(GSPC) %in% keeps2)]
VARselect(GSPC.granger.2, lag.max = 7)
grangertest(GSPC.Close.returns~USCHEX, order = 1, data=GSPC)
```


```{r}
#FOR GSPC and sentiment.LM for title first
keeps3 <- c("GSPC.Close.returns","title.sa.LM")
GSPC.granger.3 <-GSPC[ ,(names(GSPC) %in% keeps3)]
VARselect(GSPC.granger.3, lag.max = 7)
grangertest(GSPC.Close.returns~title.sa.LM, order = 5, data=GSPC)
#for desc now
keeps4 <- c("GSPC.Close.returns","desc.sa.LM")
GSPC.granger.4 <-GSPC[ ,(names(GSPC) %in% keeps4)]
VARselect(GSPC.granger.4, lag.max = 7)
grangertest(GSPC.Close.returns~desc.sa.LM, order = 5, data=GSPC)
```


```{r}
#For GSPC and sentiment.HE for title first
keeps5 <- c("GSPC.Close.returns","title.sa.HE")
GSPC.granger.5 <-GSPC[ ,(names(GSPC) %in% keeps5)]
VARselect(GSPC.granger.5, lag.max = 7)
grangertest(GSPC.Close.returns~title.sa.HE, order = 4, data=GSPC)
#for desc now
keeps6 <- c("GSPC.Close.returns","desc.sa.HE")
GSPC.granger.6 <-GSPC[ ,(names(GSPC) %in% keeps6)]
VARselect(GSPC.granger.6, lag.max = 10)
grangertest(GSPC.Close.returns~desc.sa.HE, order = 7, data=GSPC)
```

```{r}
#Now do the same for SR method
keeps7 <- c("GSPC.Close.returns","title.sr")
GSPC.granger.7 <-GSPC[ ,(names(GSPC) %in% keeps7)]
VARselect(GSPC.granger.7, lag.max = 7)
grangertest(GSPC.Close.returns~title.sr, order = 6, data=GSPC)
#for desc now
keeps8 <- c("GSPC.Close.returns","desc.sr")
GSPC.granger.8 <-GSPC[ ,(names(GSPC) %in% keeps8)]
VARselect(GSPC.granger.8, lag.max = 10)
grangertest(GSPC.Close.returns~desc.sr, order = 2, data=GSPC)
```

We have to offset our predictors as they are ordered time series. Hence, we aim to use predictors calculated at time t,to predict the direction at time t+1.
```{r}
#Remove non-relevant variables
drops <- c("GSPC.Open","GSPC.High","GSPC.Low","GSPC.Close","GSPC.Volume","GSPC.Adjusted","GSPC.Open.returns","GSPC.High.returns", "GSPC.Low.returns","GSPC.Close.returns", "GSPC.Adjusted.returns", "diff")
GSPC<-GSPC[ , !(names(GSPC) %in% drops)]
#lag predictors by 1 day
for (i in 14:45){
  GSPC[,i] <- lag(GSPC[,i],1)
}
#Remove the first observation as we now have one row of NA's
GSPC <- GSPC[-1,]
```


We cannot perform the usual resampling methods such as LOOCV or k-fold CV as our data is ordered by time. However, the caret library allows for this by using a rolling window estimation method. 
The first dataset will be the full one with all variables included

```{r}
#convert to dataframe for methods that require it
GSPC.df <- fortify.zoo(GSPC)
#gspc.clean.df$direction <- factor(as.numeric(gspc.clean.df$direction == 2))
#rownames(gspc.clean.df) <- gspc.clean.df[,1]
#gspc.clean.df <- gspc.clean.df[,-1]
levels(GSPC.df$direction) <- c("down", "up")
```

```{r}
#Split into training and test set
GSPC.end <- floor(0.8*nrow(GSPC.df)) #select the first 80% of the data
GSPC.train <- GSPC.df[1:GSPC.end,] #assign the first 80% of the data to the train set
GSPC.test <- GSPC.df[(GSPC.end+1):nrow(GSPC.df),]
```

```{r}
#Alt zoo method
GSPC.alt <- GSPC
GSPC.alt$direction <- ifelse(GSPC.alt$direction  == 1, "down", "up")
GSPC.end.alt <- floor(0.8*nrow(GSPC.alt)) #select the first 80% of the data
GSPC.train.alt <- GSPC.alt[1:GSPC.end.alt,] #assign the first 80% of the data to the train set
GSPC.test.alt <- GSPC.alt[(GSPC.end.alt+1):nrow(GSPC.alt),]
```


Create dataset with only sentiment (skip)
```{r}
GSPC.s <- GSPC.clean
dropss <- c("gspc_sentiment_title", "gspc_sentiment_desc")
GSPC.s<-GSPC.s[ , !(names(GSPC.s) %in% dropss)]
#Split into training and test set
GSPC.train.s <- GSPC.s[1:GSPC.end,] #assign the first 80% of the data to the train set
GSPC.test.s <- GSPC.s[(GSPC.end+1):nrow(GSPC.s),]
```


Prepare for caret library
```{r}
set.seed(1)
#prepare parallel computing method
registerDoParallel(cores=6)
myTimeControl <- trainControl(method = "timeslice",
                              initialWindow = 50,
                              horizon = 2,
                              fixedWindow = FALSE,
                              classProbs = TRUE, 
                              summaryFunction = twoClassSummary,
                              allowParallel = TRUE)
```


LASSO with full
```{r}
#Drop the index
GSPC.train$Index <- NULL
lasso.grid <- expand.grid(lambda=seq(0.01,0.1,length = 25),alpha=1)
lasso.full <- train(direction ~ .,
                    data = GSPC.train,
                    method = "glmnet",
                    metric = "ROC",
                    family = "binomial",
                    preProcess = c("range"),
                    trControl = myTimeControl,
                    tuneGrid = lasso.grid)
lasso.pred <- predict(lasso.full, GSPC.test)
confusionMatrix(lasso.pred, GSPC.test$direction)
#inspect model
lasso.full
plot(lasso.full)
#Find out which coefficients remained
coef(lasso.full$finalModel, lasso.full$bestTune$lambda)
```
We obtain an accuracy of 57.39%, hwoever, the model only predicted up.
The LASSO model chosen by cross validation only kept DJIA and fastK.

zoo method (skip)
```{r}
lasso.alt <- train(direction ~ .,
                    data = GSPC.train.alt,
                    method = "glmnet",
                    metric = "ROC",
                    family = "binomial",
                    preProcess = c("range"),
                    trControl = myTimeControl,
                    tuneGrid = lasso.grid)
lasso.pred.alt <- predict(lasso.alt, GSPC.test.alt)
confusionMatrix(lasso.pred.alt, GSPC.test.alt$direction)
#inspect model
lasso.alt
plot(lasso.alt)
#Find out which coefficients remained
coef(lasso.full$finalModel, lasso.full$bestTune$lambda)
```


Lasso with technical (1) only
```{r}
#Create dataset for technical indicators (1) only, without sentiment and fundamentals (skip)
GSPC.1 <- GSPC.df
technicals1 <- c("direction", "SMA10","WMA10","momentum","fastK", "fastD","stochastic.k","stochastic.d","RSI","macd","MACD", "Williams.R", "Williams.AD", "CCI")
GSPC.1<-GSPC.1[ , (names(GSPC.1) %in% technicals1)]
#Split into training and test set
GSPC.train.1 <- GSPC.1[1:GSPC.end,] #assign the first 80% of the data to the train set
GSPC.test.1 <- GSPC.1[(GSPC.end+1):nrow(GSPC.1),]
lasso.fit1 <- train(direction ~ .,
                    data = GSPC.train.1,
                    method = "glmnet",
                    metric = "ROC",
                    family = "binomial",
                    preProcess = c("range"),
                    trControl = myTimeControl,
                    tuneGrid = lasso.grid)
lasso.pred1 <- predict(lasso.fit1, GSPC.test.1)
confusionMatrix(lasso.pred1, GSPC.test.1$direction)
```
Same accuracy of 57.39%


Lasso with technical (2) only
```{r}
GSPC.2 <- GSPC.df
technicals2 <- c("direction", "OBV", "SMA5","BIAS6","PSY12","ASY1", "ASY2","ASY3","ASY4","ASY5")
GSPC.2<-GSPC.2[ , (names(GSPC.2) %in% technicals2)]
#Split into training and test set
GSPC.train.2 <- GSPC.2[1:GSPC.end,] #assign the first 80% of the data to the train set
GSPC.test.2 <- GSPC.2[(GSPC.end+1):nrow(GSPC.2),]
lasso.fit2 <- train(direction ~ .,
                    data = GSPC.train.2,
                    method = "glmnet",
                    metric = "ROC",
                    family = "binomial",
                    preProcess = c("range"),
                    trControl = myTimeControl,
                    tuneGrid = lasso.grid)
lasso.pred2 <- predict(lasso.fit2, GSPC.test.2)
confusionMatrix(lasso.pred2, GSPC.test.2$direction)
```
We obtain an accuracy of 57.39%


Lasso with fundamentals only
```{r}
GSPC.f <- GSPC.df
technicalsf <- c("direction", "USCHEX", "DJIA.returns")
GSPC.f<-GSPC.f[ , (names(GSPC.f) %in% technicalsf)]
#Split into training and test set
GSPC.train.f <- GSPC.f[1:GSPC.end,] #assign the first 80% of the data to the train set
GSPC.test.f <- GSPC.f[(GSPC.end+1):nrow(GSPC.f),]
lasso.fit.f <- train(direction ~ .,
                    data = GSPC.train.f,
                    method = "glmnet",
                    metric = "ROC",
                    family = "binomial",
                    preProcess = c("range"),
                    trControl = myTimeControl,
                    tuneGrid = lasso.grid)
lasso.pred.f <- predict(lasso.fit.f, GSPC.test.f)
confusionMatrix(lasso.pred.f, GSPC.test.f$direction)
```

LASSO with sentiment only from SentimentAnalysis method
```{r}
GSPC.sa <- GSPC.df
technicalssa <- c("direction", "title.sa.LM", "title.sa.HE", "desc.sa.LM", "desc.sa.HE")
GSPC.sa<-GSPC.sa[ , (names(GSPC.sa) %in% technicalssa)]
#Split into training and test set
GSPC.train.sa <- GSPC.sa[1:GSPC.end,] #assign the sairst 80% osa the data to the train set
GSPC.test.sa <- GSPC.sa[(GSPC.end+1):nrow(GSPC.sa),]
lasso.fit.sa <- train(direction ~ .,
                    data = GSPC.train.sa,
                    method = "glmnet",
                    metric = "ROC",
                    saamily = "binomial",
                    preProcess = c("range"),
                    trControl = myTimeControl,
                    tuneGrid = lasso.grid)
lasso.pred.sa <- predict(lasso.fit.sa, GSPC.test.sa)
confusionMatrix(lasso.pred.sa, GSPC.test.sa$direction)
```

LASSO with sentiment only from sentimentr method
```{r}
GSPC.sr <- GSPC.df
technicalssr <- c("direction", "title.sr", "desc.sr")
GSPC.sr<-GSPC.sr[ , (names(GSPC.sr) %in% technicalssr)]
#Split into training and test set
GSPC.train.sr <- GSPC.sr[1:GSPC.end,] #assign the srirst 80% osr the data to the train set
GSPC.test.sr <- GSPC.sr[(GSPC.end+1):nrow(GSPC.sr),]
lasso.fit.sr <- train(direction ~ .,
                    data = GSPC.train.sr,
                    method = "glmnet",
                    metric = "ROC",
                    sramily = "binomial",
                    preProcess = c("range"),
                    trControl = myTimeControl,
                    tuneGrid = lasso.grid)
lasso.pred.sr <- predict(lasso.fit.sr, GSPC.test.sr)
confusionMatrix(lasso.pred.sr, GSPC.test.sr$direction)
```


Random Forest
```{r}
#a quick way to find a good value for mtry
#bestMtry <- tuneRF(GSPC.train[,2:ncol(GSPC.train)], GSPC.train$direction, stepFactor = 1.5, improve = 1e-5, ntree = 1000)
#Using caret to grid search
mtry <- sqrt(ncol(GSPC.train))
rfgrid <- expand.grid(.mtry=mtry)
myGrid <- data.frame(mtry = c(2, 3, 4, 5, 6, 7, 8, 9, 10, 20, 30))
rf.fit <- train(direction ~ .,
                    data = GSPC.train,
                    method = "rf",
                    metric = 'ROC',
                    preProcess = c("range"),
                    trControl = myTimeControl,
                    tuneGrid = myGrid)
rf.test <- predict(rf.fit, newdata = GSPC.test)
confusionMatrix(rf.test, GSPC.test$direction)
rf.fit
plot(rf.fit)
```

```{r}
#Alternatively
modellist <- list()
for (ntree in c(1000, 1500, 2000, 2500)) {
	set.seed(123)
	fit <- train(direction~., data=GSPC.train, method="rf", metric="ROC", tuneGrid=myGrid, trControl=myTimeControl, ntree=ntree)
	key <- toString(ntree)
	modellist[[key]] <- fit
}
results <- resamples(modellist)
summary(results)
dotplot(results)
#Look at prediction accuracy
#first model with ntree 1000
rf.pred.1 <- predict(modellist$`1000`, GSPC.test)
confusionMatrix(rf.pred.1, GSPC.test$direction)
#second model with ntree 1500
rf.pred.2 <- predict(modellist$`1500`, GSPC.test)
confusionMatrix(rf.pred.2, GSPC.test$direction)
#third model with ntree 2000
rf.pred.3 <- predict(modellist$`2000`, GSPC.test)
confusionMatrix(rf.pred.3, GSPC.test$direction)
#fourth model with ntree 2500
rf.pred.4 <- predict(modellist$`2500`, GSPC.test)
confusionMatrix(rf.pred.4, GSPC.test$direction)
plot(modellist$`2500`)
#Variable importance
varImp(modellist$`2500`)
```
Using this method results in an increased accuracy measure for the random forest method

Now, run a randomforest with ntrees = 3000, and vary the predictors used
```{r}
rfGrid <- data.frame(mtry = c(2:13))
rf.fit.1 <- train(direction ~ .,
                    data = GSPC.train.1,
                    method = "rf",
                    metric = 'ROC',
                    preProcess = c("range"),
                    trControl = myTimeControl,
                    tuneGrid = rfGrid)
```


SVM
```{r}
#First we find the optimal tuning parameter
svm.tune <- train(direction~.,
                  data = GSPC.train,
                  method = "svmRadial",   # Radial kernel
                  tuneLength = 9,					# 9 values of the cost function
                  preProc = c("range"),  # Center and scale data
                  metric="ROC",
                  trControl=myTimeControl)
svm.tune
```

```{r}
#we narrow the grid search
svm.grid <- expand.grid(sigma = c(.020, 0.025, 0.026, 0.027, 0.028, 0.029, 0.030), C = c(0.25, 0.2, 0.35))
svm.radial <- train(direction~.,
                  data = GSPC.train,
                  method = "svmRadial",   # Radial kernel
                  tuneGrid = svm.grid,					# 9 values of the cost function
                  preProc = c("range"),  # Center and scale data
                  metric="ROC",
                  trControl=myTimeControl)
#now for poly kernel
svm.tune.poly <- train(direction~.,
                  data = GSPC.train,
                  method = "svmPoly",   # Poly kernel
                  tuneLength = 9,					# 9 values of the cost function
                  preProc = c("range"),  # Center and scale data
                  metric="ROC",
                  trControl=myTimeControl)
svm.grid.poly <- expand.grid(C = c(.020, 0.025, 0.026, 0.027, 0.028, 0.029, 0.030))
svm.poly <- train(direction~.,
                  data = GSPC.train,
                  method = "svmPoly",   # Poly kernel
                  tuneGrid = svm.grid.poly,
                  preProc = c("range"),  # Center and scale data
                  metric="ROC",
                  trControl=myTimeControl)
svm.test <- predict(svm.tune, newdata = GSPC.test)
confusionMatrix(svm.test, GSPC.test$direction)
```


Using technical indicators 1
```{r}
rfGrid <- data.frame(mtry = c(2:13))
rf.fit.1 <- train(direction ~ .,
                    data = GSPC.train.1,
                    method = "rf",
                    metric = 'ROC',
                    preProcess = c("range"),
                    trControl = myTimeControl,
                    tuneGrid = rfGrid,
                    ntree = 3000)
svm.tune.1 <- train(direction~.,
                  data = GSPC.train.1,
                  method = "svmRadial",   # Radial kernel
                  tuneLength = 9,					# 9 values of the cost function
                  preProc = c("range"),  # Center and scale data
                  metric="ROC",
                  trControl=myTimeControl)
#svm.tune.1
#we narrow the grid search
#svm.grid <- expand.grid(sigma = c(.040, .036, 0.030), C = c(0.25, 0.2, 0.35))

```



